curatedMetagenomicData

### 2019-11-11 HvL_curatedMetagenomic_NS.Rmd
updated

# the relab=FALSE is essential as it Absolute Raw Count Data
pseq = ExpressionSet2phyloseq(eset,relab=FALSE ,simplify = TRUE, phylogenetictree = TRUE)
orig_pseq <- pseq
#For the MetaPhlAn2 bugs datasets (but not other data types), you gain a lot of taxonomy-aware, ecological analysis and plotting by conversion to a phyloseq class object. curatedMetagenomicData provides the ExpressionSet2phyloseq() function to make this easy:

@@

mnd.country <- fct_collapse(mnd.country, LOW = c("NOR","SWE","FIN","DEN","LUX") , MEDIUM = c("ITA","AUT","DEU","ESP","FRA","NLD"), HIGH = c("GBR","USA","CAN") )

#mnd.country <- fct_collapse(mnd.country, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE"), HIGH = c("US") )

# reorder
mnd.country <- factor(mnd.country, levels = (c("LOW", "MEDIUM", "HIGH")))
levels(mnd.country)

@@

@@

# THis is required
mylab <- function(v1) {
  deparse(substitute(v1))
}
# aplha_d runs the diversity for each country or  groupoing and then compares the three groups
# example is alpha_d(pseq.subset,"TEST_3") this will generate a png file called TEST_3.png in the 
# currect working directory. This can be checked with getwd()
alpha_d <- function(pseq.subset,title_text) {
   # gets the meta data 
   pseq.meta <- meta(pseq.subset)
   # generates the diversites
   tab <- diversities(pseq.subset, index = "Shannon")
   # creates Shannon
   pseq.meta$Shannon <- tab$shannon
   mnd <- levels(pseq.meta$mnd) # get the variables
  # Make a pairwise list that we want to compare.
  mnd.pairs <- combn(seq_along(mnd), 2, simplify = FALSE, FUN = function(i)mnd[i])
  mnd.pairs <- list( c("LOW","MEDIUM"), c("HIGH","LOW"), c("MEDIUM","HIGH"))

   # Compare differences in Shannon index between mnd group of the study subjects
  pseq.meta$mnd <- factor(pseq.meta$mnd, levels = c("LOW","MEDIUM","HIGH"),  exclude = "NA")
  # on to the graphs!!!
  p1<-   ggviolin(pseq.meta, x = "country", y = "Shannon",
 add = "boxplot", fill = "country", palette = "paired") + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position="none") + xlab("")
# the number in the palettes must equal the number of groups in bmi.pairs

@@

# THis includes the analysis as well so can take some time to run. the graphs are generated with the label of beta diversity. It requires an input of a title. 
beta_d <- function(pseq.subset,title_text) {
  #NMDS ordination on Bray-Curtis distance
  p <<-  boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))
 # p1 = plot_ordination(pseq.subset, pseq.nmds, type = "taxa", color = "mnd", title = "Taxa by Phylum") # I dont think that we need this any more. 
  #p1 + facet_wrap(~Phylum, 3)
  pseq.nmds <- ordinate(pseq.subset, "NMDS", "bray")
  p2 <- plot_ordination(pseq.subset, pseq.nmds, type = "sample", color = "mnd")

@@

pre_perma_d <- function(pseq.subset) { 
  p.low <<- subset_samples(pseq.subset, mnd == "LOW")
  p.low.d <<- divergence(p.low)
  p.med <<- subset_samples(pseq.subset, mnd == "MEDIUM")
  p.med.d <<- divergence(p.med)
  p.high <<- subset_samples(pseq.subset, mnd == "HIGH")
  p.high.d <<- divergence(p.high)
  p <<- boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))

@@

# THis calcualtes the group PERMANOVA perma_d(p.low,p.high,"TESTHigh_vesus_Low")
perma_d <- function(x,y,title_text) { # x & y must be either p.low,p.med or p.high
  # Convert to compositional data
  pseqLvH <- merge_phyloseq(x,y) # pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p <- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = FALSE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal()  + annotate("text", x = 2.5, y = 0.02, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Beneficial Taxa")
  ggsave(paste("PERMANOVA_",title_text,".png", sep = "" ), device = "png",width = 210, height = 297, units = "mm")

@@

# comp_d <- function(pseq,subset,title_text) {
#   png(paste(title_text,"Phylum_abundance.png",sep="_"))
#   p <- plot_bar(pseq.subset,"Phylum") + theme(axis.text.x = element_text(face="italic"))
#   print(p)

#   prev.otu <- plot_taxa_prevalence(pseq.subset, "Phylum")
#   prev.otu + theme(legend.position="none")
#   ggsave(paste(title_text,"phylum_prevalence.png",sep="_"))
#   tab <- plot_taxa_prevalence(pseq.subset, "Phylum", detection = 5)
#   tab + theme(legend.position="none")
#   ggsave(paste(title_text,"Phylum_prevalence_detection.png",sep="_"))

#   pseq.ph <- aggregate_taxa(pseq.subset, "Phylum")
#   pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")
#   plot.comp.rel <- plot_composition(pseq.ph.rel, x.label = "sample") +
#     theme(legend.position = "bottom") + theme_bw() +
#     theme(axis.text.x=element_blank()) +
#     c + scale_fill_brewer(palette = "Paired") + guides(fill=guide_legend("Phylum"))
#   print(plot.comp.rel)
#   ggsave("phylum_relative_abundance11.png")
#   plot.comp.rel + theme(legend.position="none")
#   ggsave("phylum_relative_abundance12.png")

#   pseq.ph <- aggregate_taxa(pseq, "Phylum")
#   pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")
#   plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND   Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))
#   plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
#   ggsave("phylum_abundance_bar.png")
#   # Top 5
#   top5P.names = sort(tapply(taxa_sums(pseq), tax_table(pseq)[, "Genus"], sum), TRUE)[1:5]

#   top5P = subset_taxa(pseq, Genus %in% names(top5P.names))
#   pseq.ph.rel <- microbiome::transform(top5P, "compositional")
#   plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))

#   plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma")

#   # Top 10
#   top10P.names = sort(tapply(taxa_sums(pseq), tax_table(pseq)[, "Genus"], sum), TRUE)[1:10]

#   top10P = subset_taxa(pseq, Genus %in% names(top10P.names))
#   pseq.ph.rel <- microbiome::transform(top10P, "compositional")

#   plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))

#   plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
#   # new
#   p.new <- aggregate_taxa(pseq, "Phylum")
#   p.new.rel <- microbiome::transform(p.new, "compositional")

#   plot.p.new.rel <- plot_composition(p.new.rel, average_by = "mnd", x.label = " ", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum")) + theme(legend.text = element_text( face="italic"))

#   plot.p.new.rel + scale_fill_brewer(palette = "Paired") 
#   ggsave("pnew_abundance_bar.png")

#   # Top Phylum Heatmap
#   pseq.ph <- aggregate_taxa(pseq, "Phylum")

#   pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")

#   plot.comp.rel <- plot_composition(pseq.ph.rel, x.label = "sample", plot.type = "heatmap")

#   plot(plot.comp.rel+ coord_flip()) + theme(axis.text.x=element_blank())
#   ggsave("phylum_histogram.png")

#   #p.new heatmap
#   pseq.n.ph <- aggregate_taxa(pseq, "Phylum")
#   pseq.n.ph.rel <- microbiome::transform(pseq.n.ph, "compositional")

#   plot.n.comp.rel <- plot_composition(pseq.n.ph.rel, x.label = "sample", plot.type = "heatmap")

#   p <- plot.n.comp.rel + coord_flip()
#   p + theme(axis.text.x=element_blank()) + labs(y = "Sample")
#   ggsave("pnew_histogram.png")

#   # Newwork
#   png("network_bray_phylum.png")
#   jg = make_network(pseq, "taxa", "bray", 0.3)
#   plot_network(jg, pseq, "taxa", color = "Phylum", line_weight = 0.4, label = NULL)
#   dev.off()
#   #Community-composition-plots
#   # Create a data table for ggplot
# # agglomerate at phylum level
# pseq_phy <- tax_glom(pseq.new, taxrank = "Phylum")

# # Transform to rel. abundance (or use ps0.ra)
# pseq_phyl <- transform_sample_counts(pseq_phy, function(x) {x/sum(x)})

# # Melt to long format for easy ggploting
# pseq_phylum <- psmelt(pseq_phyl)

#      # filter(Abundance > 0.01)                             # Filter out low abundance taxa
# # Plot - Phylum
#   p.ra.phylum <- ggplot(pseq_phylum, aes(x = sample, y = Abundance, fill = Phylum)) + 
#   geom_bar(stat = "identity", width = 1) +
#   facet_wrap(~mnd, scales = "free_x", nrow = 4, ncol = 7) +
#   theme(axis.text.x = element_blank()) +
#   theme(axis.title.x = element_blank()) +
#   labs(title = "Abundant Phylum (> 1%)")

#   p <- p.ra.phylum + scale_fill_viridis(discrete = TRUE, option = "A")
#   plot(p)
#   ggsave("phylum_abun_mnd.png")
# }

@@

alpha_d(pseq.subset,"TEST_3")
beta_d(pseq.subset,"TEST_2")
pre_perma_d(pseq.subset) 
perma_d(p.low,p.high,"TESt_highVlow")

>>>>

# the relab=FALSE is essential as it Absolute Raw Count Data
pseq = ExpressionSet2phyloseq(eset,relab=FALSE ,simplify = TRUE, phylogenetictree = TRUE)
orig_pseq <- pseq
#For the MetaPhlAn2 bugs datasets (but not other data types), you gain a lot of taxonomy-aware, ecological analysis and plotting by conversion to a phyloseq class object. curatedMetagenomicData provides the ExpressionSet2phyloseq() function to make this easy:
data(atlas1006)
atlas.pseq <- baseline(atlas1006)

@@

# Cut off is 1.8  Low

mnd.country <- fct_collapse(mnd.country, LOW = c("DEU","ESP","ITA","AUT","DNK","LUX" ) , HIGH = c("NOR","FIN","FRA","NLD","GBR","USA","CAN","SWE") )

table(mnd.country)
#mnd.country <- fct_collapse(mnd.country, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE"), HIGH = c("US") ) "ITA","AUT","DEU","ESP","FRA","NLD"

# reorder
mnd.country <- factor(mnd.country, levels = (c("LOW", "HIGH")))
levels(mnd.country)

@@

# THIS IS FOR THE ATLAS 1006 
# Create mnd variable
# grabs the nationality from phyloseq
mnd.nationality <- get_variable(atlas.pseq, "nationality" )

# this collapses the two varialbes into a new variable AB in this case. The command for multiple changes is fct_collapse(x, AB = c("A","B"), DE = c("D","E"))
mnd.nationality <- fct_collapse(mnd.nationality, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE"), HIGH = c("US") )

# reorder
mnd.nationality <- factor(mnd.nationality, levels = (c("LOW", "MEDIUM", "HIGH")))
levels(mnd.nationality)

# creates a new variable in the phyloseq called mnd
sample_data(atlas.pseq)$mnd = mnd.nationality

# checks that it has worked.
get_variable(atlas.pseq, "mnd")

@@

# THis is required
mylab <- function(v1) {
  deparse(substitute(v1))
}
# aplha_d runs the diversity for each country or  groupoing and then compares the three groups
# example is alpha_d(pseq.subset,"TEST_3") this will generate a png file called TEST_3.png in the 
# currect working directory. This can be checked with getwd()
alpha_d_two <- function(pseq.subset,title_text) {
   # gets the meta data 
   pseq.meta <- meta(pseq.subset)
   # generates the diversites
   tab <- microbiome::alpha(pseq.subset, index = "shannon")

   # creates Shannon
   pseq.meta$Shannon <- tab$diversity_shannon
   mnd <- levels(pseq.meta$mnd) # get the variables
  # Make a pairwise list that we want to compare.
  mnd.pairs <- combn(seq_along(mnd), 2, simplify = FALSE, FUN = function(i)mnd[i])
  mnd.pairs <- list( c("HIGH","LOW"))

   # Compare differences in Shannon index between mnd group of the study subjects
  pseq.meta$mnd <- factor(pseq.meta$mnd, levels = c("LOW","HIGH"),  exclude = "NA")
  # on to the graphs!!! country 
  p1<-   ggviolin(pseq.meta, x = "nationality", y = "Shannon",
 add = "boxplot", fill = "nationality", palette = "paired") + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position="none") + xlab("")
# the number in the palettes must equal the number of groups in bmi.pairs

@@

# THis includes the analysis as well so can take some time to run. the graphs are generated with the label of beta diversity. It requires an input of a title. 
beta_d_two <- function(pseq.subset,title_text) {
  #NMDS ordination on Bray-Curtis distance
  p <<-  boxplot(list(Low = p.low.d, High = p.high.d))
 # p1 = plot_ordination(pseq.subset, pseq.nmds, type = "taxa", color = "mnd", title = "Taxa by Phylum") # I dont think that we need this any more. 
  #p1 + facet_wrap(~Phylum, 3)
  pseq.nmds <- ordinate(pseq.subset, "NMDS", "bray")
  p2 <- plot_ordination(pseq.subset, pseq.nmds, type = "sample", color = "mnd")

@@

pre_perma_d_two <- function(pseq.subset) { 
  p.low <<- subset_samples(pseq.subset, mnd == "LOW")
  p.low.d <<- divergence(p.low)
  p.high <<- subset_samples(pseq.subset, mnd == "HIGH")
  p.high.d <<- divergence(p.high)
  p <<- boxplot(list(Low = p.low.d, High = p.high.d))

@@

# THis calcualtes the group PERMANOVA perma_d(p.low,p.high,"TESTHigh_vesus_Low")
perma_d_two <- function(title_text) { # x & y must be either p.low,p.med or p.high
  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
  pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p <- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = FALSE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 20)  + annotate("text", x = 2.5, y = 0.02, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Beneficial Taxa")  
  ggsave(paste("PERMANOVA_",title_text,".png", sep = "" ), device = "png",width = 210, height = 297, units = "mm")
  ggsave("test_permanova.png", device = "png",width = 297, height = 210, units = "mm")
 }

testpseq <- atlas1006 %>%
   subset_samples(DNA_extraction_method == "r") %>%
   aggregate_taxa(level = "Phylum") %>%
   transform(transform = "compositional")
p <- plot_composition(testpseq, sample.sort = "Firmicutes",
         otu.sort = "abundance", verbose = TRUE) +
     scale_fill_manual(values = default_colors("Phylum")[taxa(testpseq)])

comp_d <- function(pseq,subset,title_text) {
  png(paste(title_text,"Phylum_abundance.png",sep="_"))
  p <- plot_bar(pseq.subset,"Phylum") + theme(axis.text.x = element_text(face="italic"))
  print(p)

  prev.otu <- plot_taxa_prevalence(pseq.subset, "Phylum" , detection = 5)
  prev.otu + theme(legend.position="none")
  ggsave(paste(title_text,"phylum_prevalence.png",sep="_"))
  tab <- plot_taxa_prevalence(pseq.subset, "Phylum", detection = 5)
  tab + theme(legend.position="none")
  ggsave(paste(title_text,"Phylum_prevalence_detection.png",sep="_"))

  pseq.ph <- aggregate_taxa(pseq.subset, "Phylum")
  pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")
  plot.comp.rel <- plot_composition(pseq.ph.rel, x.label = "sample") +
    theme(legend.position = "bottom") + theme_bw() +
    theme(axis.text.x=element_blank()) +
    c + scale_fill_brewer(palette = "Paired") + guides(fill=guide_legend("Phylum"))
  print(plot.comp.rel)
  ggsave("phylum_relative_abundance11.png")
  plot.comp.rel + theme(legend.position="none")
  ggsave("phylum_relative_abundance12.png")

  pseq.ph <- aggregate_taxa(pseq.subset, "Phylum")
  pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")
  plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND   Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))
  plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
  ggsave("phylum_abundance_bar.png")
  # Top 5
  top5P.names = sort(tapply(taxa_sums(pseq), tax_table(pseq)[, "Genus"], sum), TRUE)[1:5]

  top5P = subset_taxa(pseq, Genus %in% names(top5P.names))
u  pseq.ph.rel <- microbiome::transform(top5P, "compositional")
  plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))

  plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma")

  # Top 10
  top10P.names = sort(tapply(taxa_sums(pseq), tax_table(pseq)[, "Genus"], sum), TRUE)[1:10]

  top10P = subset_taxa(pseq, Genus %in% names(top10P.names))
  pseq.ph.rel <- microbiome::transform(top10P, "compositional")

  plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))

  plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
  # new
  p.new <- aggregate_taxa(pseq, "Phylum")
  p.new.rel <- microbiome::transform(p.new, "compositional")

  plot.p.new.rel <- plot_composition(p.new.rel, average_by = "mnd", x.label = " ", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum")) + theme(legend.text = element_text( face="italic"))

  plot.p.new.rel + scale_fill_brewer(palette = "Paired") 
  ggsave("pnew_abundance_bar.png")

  # Top Phylum Heatmap
  pseq.ph <- aggregate_taxa(pseq, "Phylum")

  pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")

  plot.comp.rel <- plot_composition(pseq.ph.rel, x.label = "sample", plot.type = "heatmap")

  plot(plot.comp.rel+ coord_flip()) + theme(axis.text.x=element_blank())
  ggsave("phylum_histogram.png")

  #p.new heatmap
  pseq.n.ph <- aggregate_taxa(pseq, "Phylum")
  pseq.n.ph.rel <- microbiome::transform(pseq.n.ph, "compositional")

  plot.n.comp.rel <- plot_composition(pseq.n.ph.rel, x.label = "sample", plot.type = "heatmap")

  p <- plot.n.comp.rel + coord_flip()
  p + theme(axis.text.x=element_blank()) + labs(y = "Sample")
  ggsave("pnew_histogram.png")

  # Newwork
  png("network_bray_phylum.png")
  jg = make_network(pseq, "taxa", "bray", 0.3)
  plot_network(jg, pseq, "taxa", color = "Phylum", line_weight = 0.4, label = NULL)
  dev.off()
  #Community-composition-plots
  # Create a data table for ggplot
# agglomerate at phylum level
pseq_phy <- tax_glom(pseq.subset, taxrank = "Phylum")

# Transform to rel. abundance (or use ps0.ra)
pseq_phyl <- transform_sample_counts(pseq_phy, function(x) {x/sum(x)})

# Melt to long format for easy ggploting
pseq_phylum <- psmelt(pseq_phyl)

     # filter(Abundance > 0.01)                             # Filter out low abundance taxa
# Plot - Phylum
    p.ra.phylum <- ggplot(pseq_phylum, aes(x = sample, y = Abundance, fill = Phylum)) + 
    geom_bar(stat = "identity", width = 1) +
    facet_wrap(~mnd, scales = "free_x", nrow = 4, ncol = 7) +
    theme(axis.text.x = element_blank()) +
    theme(axis.title.x = element_blank()) +
    labs(title = "Abundant Phylum (> 1%)")

  p <- p.ra.phylum + scale_fill_viridis(discrete = TRUE, option = "A")
  plot(p)
  ggsave("phylum_abun_mnd.png")

 # these are other functions required along the way.

    # transform counts to n
  physeq.scale <- transform_sample_counts(physeq, 
    function(x) {(n * x/sum(x))}
  )

  # Pick the rounding functions
  if (round == "floor"){
    otu_table(physeq.scale) <- floor(otu_table(physeq.scale))
  } else if (round == "round"){
    otu_table(physeq.scale) <- myround(otu_table(physeq.scale))
  }

  # Prune taxa and return new phyloseq object
  physeq.scale <- prune_taxa(taxa_sums(physeq.scale) > 0, physeq.scale)
  return(physeq.scale)
}

##### ADONIS ###########

# Function to run adonis test on a phyloseq object and a variable from metadata
# Make sure OTU data is standardized/normalized before 
phyloseq_to_adonis <- function(physeq, distmat = NULL, dist = "bray", formula) {

  if(!is.null(distmat)) {
    phydist <- distmat
  } else {
    phydist <- phyloseq::distance(physeq, dist)
  }

  metadata <- as(sample_data(physeq), "data.frame")

  # Adonis test
  f <- reformulate(formula, response = "phydist")
  adonis.test <- adonis(f, data = metadata)
  print(adonis.test)

  # Run homogeneity of dispersion test if there is only 1 variable
  if (grepl("\\+", formula)) {
    l <- list(
      dist = phydist, 
      formula = f, 
      adonis = adonis.test
    )
  } else {
    group <- metadata[ ,formula]
    beta <- betadisper(phydist, group)
    disper.test = permutest(beta)
    print(disper.test)

    l <- list(
      dist = phydist, 
      formula = f, 
      adonis = adonis.test, 
      disper = disper.test
    )
  }
  return (l)
}

########## Bar Plots #################

# This function takes a phyloseq object, agglomerates OTUs to the desired taxonomic rank, 
# prunes out OTUs below a certain relative proportion in a sample (ie 1% ) 
# and melts the phyloseq object into long format which is suitable for ggplot stacked barplots.
taxglom_and_melt <- function(physeq, taxrank, prune){

  # Agglomerate all otu's by given taxonomic level
  pglom <- tax_glom(physeq, taxrank = taxrank)

  # Create a new phyloseq object which removes taxa from each sample below the prune parameter
  pglom_prune <- transform_sample_counts(pglom,function(x) {x/sum(x)})
  otu_table(pglom_prune)[otu_table(pglom_prune) < prune] <- 0
  pglom_prune <- prune_taxa(taxa_sums(pglom_prune) > 0, pglom_prune)

  # Melt into long format and sort by taxonomy
  physeq_long <- psmelt(pglom_prune)
  physeq_long <- physeq_long[order(physeq_long[ ,taxrank]), ]

  # Return long data frame
  return(physeq_long)
}

###### Merge functions ############

# Merge samples by averaging OTU counts instead of summing
merge_samples_mean <- function(physeq, group, round){
  # Calculate the number of samples in each group
  group_sums <- as.matrix(table(sample_data(physeq)[ ,group]))[,1]

  # Merge samples by summing
  merged <- merge_samples(physeq, group)

  # Divide summed OTU counts by number of samples in each group to get mean
  # Calculation is done while taxa are columns
  x <- as.matrix(otu_table(merged))
  if(taxa_are_rows(merged)){ x<-t(x) }

  # Pick the rounding functions
  if (round == "floor"){
    out <- floor(t(x/group_sums))
  } else if (round == "round"){
    out <- myround(t(x/group_sums))
  }

  # Return new phyloseq object with taxa as rows
  out <- otu_table(out, taxa_are_rows = TRUE)
  otu_table(merged) <- out
  return(merged)
}

# Merge samples, just including OTUs that were present in all merged samples
# Call this function before running merge_samples()
merge_OTU_intersect <- function(physeq, group){

  # Make sure we're not starting with more taxa than we need 
  physeq <- prune_taxa(taxa_sums(physeq) > 0, physeq)

  s <- data.frame(sample_data(physeq))
  l <- levels(s[,group])
  o <- otu_table(physeq)

  # Loop through each category
  for (cat in 1:length(l)) {

    # Get the index of all samples in that category
    w <- which(s[,group]==l[cat])

    # subset to just those columns of OTU table
    cat.sub<-o[,w]
    print(dim(cat.sub))

    # Find the indices of 0's in the OTU table
    zeros <- apply(cat.sub, 1, function(r) any(r == 0))

    # If an OTU had a 0 in at least one sample, change all samples to 0
    cat.sub[zeros,] <- 0
  }

  o[,w] <- cat.sub
  otu_table(physeq) <- o

  return(physeq)

@@

@@

alpha_d_two(pseq.subset,"cHL")
beta_d_two(pseq.subset,"cHL")
alpha_d_two(atlas.pseq,"aHL")
beta_d_two(atlas.pseq,"aHL")

pre_perma_d_two(atlas.pseq) 
perma_d_two("1303_highVlow")
pre_perma_d_two(pseq.subset) 
perma_d_two("1253_highVlow")
